### 2020-08-16 subgroup_analysis.Rmd
Update repo

---
title: "Subgroup cost analysis"
output: html_notebook
---

```{r setup, echo = FALSE}
setwd("~/interval-sims/src")
library(knitr)
library(dplyr)
library(plyr)
library(lubridate)
library(numbers)
library(data.table)
library(R.utils)
source("pffunctions.R")
source("helper_functions.R")
```

```{r load_data, echo = FALSE}
load("~/eba_sims/data/20191211CollectionAmounts.rdata")
```

# Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")

# Women under 30
fU30Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Women", age = c(17, 30), Hb_deferral = 1)
fU30don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Women", status = "-", age = c(17, 30))

# Women under 40
fU40Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Women", age = c(30, 40), Hb_deferral = 1)
fU40don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Women", status = "-", age = c(30, 40))

# Men under 40
mU40Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Men", age = c(17, 40), Hb_deferral = 1)
mU40don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Men", status = "-", age = c(17, 40))
```

We don't need to analyse the entire available history to get a robust average. It is also important that the statistics reflect the current state of the donation population. This is why we cut the sets into windows.

## Windowing
```{r windowing}
# Windowing between 2018-2020
wfU30don <- subset(fU30don, date >= "2018-01-01" & date <= "2019-12-30")
wfU30def <- subset(fU30Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")

wfU40don <- subset(fU40don, date >= "2018-01-01" & date <= "2019-12-30")
wfU40def <- subset(fU40Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")

wmU40don <- subset(mU40don, date >= "2018-01-01" & date <= "2019-12-30")
wmU40def <- subset(mU40Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")
```

# Strategy 1

In the first strategy we limit the maximum donations for women under 30 to 1 per year. We thus iterate through the 2 year window, identify donors and see how many times they've donated and how many times they've been deferred to calculate costs and savings with this strategy. We limit the analysis to first 10,000 events to reduce the analysis time.

```{r tallying1}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wfU30don$donor[i]
  dondate <- as.Date(head(wfU30don$date[i], 10))

  # Get donor window
  yearw <- subset(wfU30don, date >= dondate & date <= (dondate + years(1)))
  defyearw <- subset(wfU30def, date >= dondate & date <= (dondate + years(1)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}
```

We also need to check what percentage of all donations come from this subgroup
```{r total_proportion}
wdons <- subset(dons, date >= "2018-01-01" & date <= "2019-12-30")
p <- nrow(wfU30don)/nrow(wdons)
p
```

## Cost effect

Cost effect is calculated by multiplying the lost flux by the unit marketing cost (representing increasing marketing costs) and then subtracting savings generated by avoided deferrals from it. Deferrals are scaled w.r.t total donation flux (max avoided is around 3 %).

```{r dons_lost1}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided1}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect1}
E <- 2*dons_lost - 60*def_avoided
E
```

# Strategy 2

Second strategy limits donations from women under 30 years to 1 per year and to 2 per year for women between 30-40. We'll use the result of the previous strategy and add to it results from limiting women between 30-40 to 2 donations a year.

```{r tallying2}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wfU40don$donor[i]
  dondate <- as.Date(head(wfU40don$date[i], 10))

  # Get donor window
  yearw <- subset(wfU40don, date >= dondate & date <= (dondate + months(6)))
  defyearw <- subset(wfU40def, date >= dondate & date <= (dondate + months(6)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}
```

We also need to check what percentage of all donations come from this subgroup
```{r total_proportion2}
p <- nrow(wfU40don)/nrow(wdons)
p
```

## Cost effect

```{r dons_lost2}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided2}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect2}
E2 <- 2*dons_lost - 60*def_avoided
E12 <- E + E2
E12
```

# Strategy 3

Under the 3rd strategy, women under 30 may also donate twice per year. We recalulate and sum with E2.

```{r tallying3}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wfU30don$donor[i]
  dondate <- as.Date(head(wfU30don$date[i], 10))

  # Get donor window
  yearw <- subset(wfU30don, date >= dondate & date <= (dondate + months(6)))
  defyearw <- subset(wfU30def, date >= dondate & date <= (dondate + months(6)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}

# Reinstance p
p <- nrow(wfU30don)/nrow(wdons)
p
```

## Cost effect

```{r dons_lost3}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided3}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect3}
E3 <- 2*dons_lost - 60*def_avoided
E23 <- E2 + E3
E23
```

# Strategy 4

4 donations for men under 40.

```{r tallying4}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wmU40don$donor[i]
  dondate <- as.Date(head(wmU40don$date[i], 10))

  # Get donor window
  yearw <- subset(wmU40don, date >= dondate & date <= (dondate + months(3)))
  defyearw <- subset(wmU40def, date >= dondate & date <= (dondate + months(3)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}

# Reinstance p
p <- nrow(wmU40don)/nrow(wdons)
p
```

## Cost effect

```{r dons_lost4}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided4}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect4}
E4 <- 2*dons_lost - 60*def_avoided
E4
```