### 2020-08-23 src/eba_sims_w_data.Rmd
Add files via upload

---
title: "eba_sims_data"
author: "Esa Turkulainen"
date: "1/13/2020"
output: pdf_document
---

```{r setup, echo = FALSE}
setwd("~/interval-sims/src")
library(ggplot2)
library(plotly)
library(gridExtra)
library(knitr)
library(dplyr)
library(plyr)
library(lubridate)
library(numbers)
library(data.table)
library(R.utils)
library(forecast)
source("pffunctions.R")
source("helper_functions.R")
```

```{r load_data, echo = FALSE}
load("~/eba_sims/data/20191211CollectionAmounts.rdata")
```

## Inspect data
```{r inspect}
summary(donation10)
```

```{r filter_test}
# Only donat_phleb="K" and status="-"
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
```

```{r distribution_of_event_counts}
eventcounts <- table(dons$donor)
hist(eventcounts)
```
Sanity check: passed. (60 is the maximum number of donations in 10 years time)

```{r}
# Order donations by date
dons <- dons[with(dons, order(date)), ]

# Transform dates to yearweeks
dons$date <- yearweek(dons$date)

# Extract first time donations
firsts <- dons[!duplicated(dons$donor),]
```

```{r}
# Firsts
first_count <- as.data.frame(table(factor(firsts$date, levels=unique(firsts$date))))
# (We needed to call factor() to retain the chronological order of the rows)
colnames(first_count) <- c("week", "firsts")

# All
all_count <- as.data.frame(table(factor(dons$date, levels=unique(dons$date))))
colnames(all_count) <- c("week", "all")
```

```{r}
# Combine into one neat df
counts_df <- data.frame(week=first_count$week, 
                        first=first_count$firsts, 
                        all=all_count$all)
```

```{r}
# Plot
ggplot(data = counts_df, aes(x=week, y=first)) + 
  geom_point(color = "dark red") +
  geom_point(aes(x=week, y=all), color = "pink") +
  theme(axis.text.x = element_text(angle = 45)) +
  geom_abline(intercept = mean(counts_df[158:520,2]), slope=0) +
  geom_abline(intercept = mean(counts_df[158:520,3]), slope=0) +
  scale_x_discrete(breaks=c("2010.1", "2011.1", "2012.1", "2013.1", "2014.1", "2015.1", "2016.1", "2017.1", "2018.1", "2019.1")) +
  theme_bw() +
  labs(title = "Number of weekly first time donations vs. all donations",
       subtitle = paste0("Each week around ", round(mean(counts_df[158:520,2])/mean(counts_df[158:520,3])*100, 2), "%", " of all donations comprise first time donors (2013 onwards)."),
       x = "Weeks",
       y = "Count") +
  annotate(geom="text", 
           x="2019.1", 
           y=mean(counts_df[158:520,2])+400, 
           label=paste0("Firsts: ", round(mean(counts_df[158:520,2]), 1))) +
  annotate(geom="text",
           x="2019.1",
           y=mean(counts_df[158:520,3])+400, 
           label=paste0("All: ", round(mean(counts_df[158:520,3]), 1)))
```

```{r}
library(truncnorm)

q <- seq(0, 0.031, 0.031/1000)
x <- seq(1, 1.5, 1/10000)
adjustments <- sort(round(rtruncnorm(n=length(q), a=1, b=1.5, mean=1.1, sd=0.1), digits = 4))
E.matrix <- matrix(, nrow = 1001, ncol = 1001)
Pm <- 2
Pd <- 60 # from Satu
rl <- 0
Fn = 0.259 # Vuosikertomus 2019
qalt <- seq(0, 0.06, 0.06/1000) # Larger q
q <- seq(0, 0.031, 0.031/1000) # from data
for(i in 1:1001){
  Fratio <- adjustments[i] / (1 + Fn * (adjustments[i] - 1))
  for(j in 1:1001){
    Em <- Pm * (Fratio - 1 - q[j] * rl)
    Ed <- Pd * q[j]
    E.matrix[i, j] <- Em - Ed
    }
}

E.long <- melt(E.matrix)

```

```{r}
colnames(E.long) <- c("adjustment", "q", "cost")
ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
geom_tile(aes(fill=cost)) +
  stat_contour(color = "white", size = 0.7, bins = 2) +
  scale_fill_gradientn(colours=c("blue", "red")) +
labs(x="q%", y="Adjustment", title="Cost surface", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ". Cost effect thresholds highlighted at 0 and -1 euros.")) +
theme_bw() +
scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.1")) +
scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
```