### 2020-09-21 src/current_model_effect.Rmd
Update to most recent

---
title: "Current model cost effect"
author: "Esa Turkulainen"
date: "8/25/2020"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
setwd("~/interval-sims/src")
library(knitr)
library(dplyr)
library(plyr)
library(lubridate)
library(numbers)
library(data.table)
library(R.utils)
source("pffunctions.R")
source("helper_functions.R")
```

```{r load_data, echo = FALSE}
load("~/eba_sims/data/20191211CollectionAmounts.rdata")
```

# Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
# Order donations by date
dons <- dons[with(dons, order(date)), ]
# Find only returning donors
rdons <- dons[duplicated(dons$donor),]

# Successful male
mdon <- filter_donations(data = rdons, donat_phleb = "K", status = "-", gender = "Men")
# Successful female
fdon <- filter_donations(data = rdons, donat_phleb = "K", status = "-", gender = "Women")
# Deferrals
def <- filter_donations(data = donation10, donat_phleb = "*", Hb_deferral = 1)
mdef <- filter_donations(data = def, gender = "Men")
fdef <- filter_donations(data = def, gender = "Women")

# Windows
walldons <- subset(dons, date >= "2018-01-01" & date <= "2019-12-30")
wdons <- subset(rdons, date >= "2018-01-01" & date <= "2019-12-30")
wmdon <- subset(mdon, date >= "2018-01-01" & date <= "2019-12-30")
wfdon <- subset(fdon, date >= "2018-01-01" & date <= "2019-12-30")
wdef <- subset(def, date >= "2018-01-01" & date <= "2019-12-30")
wmdef <- subset(mdef, date >= "2018-01-01" & date <= "2019-12-30")
wfdef <- subset(fdef, date >= "2018-01-01" & date <= "2019-12-30")
```

# Find ratios
```{r}
# Male ratio
mr <- nrow(wmdon)/nrow(wdons)
# Female ratio
fr <- nrow(wfdon)/nrow(wdons)
# Male deferral ratio
mdr <- nrow(wmdef) / nrow(wdef)
# Female deferral ratio
fdr <- nrow(wfdef) / nrow(wdef)
# Total deferral prevalence
d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
```

The current model finds 1653/2116 = 78.1% of deferrals. In addition, it misclassifies 22.57% of donations as deferrals.
```{r}
TPR <- 0.781
FPR <- 0.2257
# Find q
q <- d * TPR
# Find a for 6 month category
# correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
# incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
a6 <- 1 * ((1 - FPR) * (1 - d)) + 
  3 * mr * (FPR * (1 - d)) + 2 * fr * (FPR * (1 - d)) + 
  3 * mdr * d + 2 * fdr * d

a12 <- 1 * ((1 - FPR) * (1 - d)) + 
  6 * mr * (FPR * (1 - d)) + 4 * fr * (FPR * (1 - d)) + 
  6 * mdr * d + 4 * fdr * d
```

# Final cost
```{r}
Pm <- 2
Pd <- 60 # from Satu
rl <- 0
Fn = 0.1066 # Data

# a6
Fratio <- a6 / (1 + Fn * (a6 - 1))
Em <- Pm * (Fratio - 1 - q * rl)
Ed <- Pd * q
E6 <- Em - Ed

# a12
Fratio <- a12 / (1 + Fn * (a12 - 1))
Em <- Pm * (Fratio - 1 - q * rl)
Ed <- Pd * q
E12 <- Em - Ed
```

```{r}
E6
E12
```