### 2020-09-21 src/current_model_effect.Rmd
cleaning and testing variable probability cut-off

---
title: "Current model cost effect"
author: "Esa Turkulainen"
date: "8/25/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
#setwd("~/interval-sims/src")
library(tidyverse)
library(truncnorm)
library(caret)
source("helper_functions.R")
```

```{r load_data, echo = FALSE}
load("..//data/20191211CollectionAmounts.rdata")
```

The data object donation10 contains donation history data from beginning of 2010 to end of 2019 it has following columns.

```{r }
lapply(donation10,class)
```

# Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
# Order donations by date
dons <- dons[with(dons, order(date)), ]
# Find only returning donors
rdons <- dons[duplicated(dons$donor),]

# Successful male
mdon <- filter_donations(data = rdons, donat_phleb = "K", status = "-", gender = "Men")
# Successful female
fdon <- filter_donations(data = rdons, donat_phleb = "K", status = "-", gender = "Women")
# Deferrals
def <- filter_donations(data = donation10, donat_phleb = "*", Hb_deferral = 1)
mdef <- filter_donations(data = def, gender = "Men")
fdef <- filter_donations(data = def, gender = "Women")

# Windows
walldons <- subset(dons, date >= "2018-01-01" & date <= "2019-12-30")
wdons <- subset(rdons, date >= "2018-01-01" & date <= "2019-12-30")
wmdon <- subset(mdon, date >= "2018-01-01" & date <= "2019-12-30")
wfdon <- subset(fdon, date >= "2018-01-01" & date <= "2019-12-30")
wdef <- subset(def, date >= "2018-01-01" & date <= "2019-12-30")
wmdef <- subset(mdef, date >= "2018-01-01" & date <= "2019-12-30")
wfdef <- subset(fdef, date >= "2018-01-01" & date <= "2019-12-30")
```

# Find ratios
```{r}
# Male ratio
mr <- nrow(wmdon)/nrow(wdons)
# Female ratio
fr <- nrow(wfdon)/nrow(wdons)
# Male deferral ratio
mdr <- nrow(wmdef) / nrow(wdef)
# Female deferral ratio
fdr <- nrow(wfdef) / nrow(wdef)
# Total deferral prevalence
d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
```

The current model finds 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 21.45% of donations as deferrals.
```{r}
TPR <- 1662/2116
FPR <- 1- (1662 /2116)
# Find q
q <- d * TPR
# Find a for 6 month category
# correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
# incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
a6 <- 1 * ((1 - FPR) * (1 - d)) + 
  3 * mr * (FPR * (1 - d)) + 2 * fr * (FPR * (1 - d)) + 
  3 * mdr * d + 2 * fdr * d

a12 <- 1 * ((1 - FPR) * (1 - d)) + 
  6 * mr * (FPR * (1 - d)) + 4 * fr * (FPR * (1 - d)) + 
  6 * mdr * d + 4 * fdr * d
TPR
FPR
a6
a12
```

# Final cost
```{r}
Pm <- 2 # estimate collected from FRCBS experts
Pd <- 60 # estimate collected from FRCBS experts
rl <- 0 # estimated previously from data
Fn = 0.1066 # etimated from data in https://github.com/FRCBS/interval-sims/blob/master/src/eba_sims_w_data.Rmd

# a6
Fratio <- a6 / (1 + Fn * (a6 - 1))
Em <- Pm * (Fratio - 1 - q * rl)
Ed <- Pd * q
E6 <- Em - Ed

# a12
Fratio <- a12 / (1 + Fn * (a12 - 1))
Em <- Pm * (Fratio - 1 - q * rl)
Ed <- Pd * q
E12 <- Em - Ed
```

```{r}
E6
E12
q
```

# Plot final cost on data surface

```{r}

q.seq <- seq(0, 0.032, 0.032/1000)
adjustments <- seq(1, 2, 1/1000)
E.matrix <- matrix(nrow = 1001, ncol = 1001)
Fn = 0.1066 # 
for(i in 1:1001){
  Fratio <- adjustments[i] / (1 + Fn * (adjustments[i] - 1))
  for(j in 1:1001){
    Em <- Pm * (Fratio - 1 - q.seq[j] * rl)
    Ed <- Pd * q.seq[j]
    E.matrix[i, j] <- Em - Ed
    }
}

E.long <- reshape2::melt(E.matrix)

```

```{r}
colnames(E.long) <- c("adjustment", "q", "cost")
ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
geom_tile(aes(fill=cost)) +
  stat_contour(color = "white", size = 0.7, bins = 2) +
  scale_fill_gradientn(colours=c("blue", "red")) +
labs(x="q%", y="Adjustment", title="Cost surface (2018-2020)", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ". Cost effect thresholds highlighted at 0 and -1 euros.", "\n", "Dots mark current model performance with a 6 or 12 month extension strategy.")) + 
  annotate("point", x = head(which(q.seq > q ),1), y = which(adjustments > a6 & adjustments < (a6 + 0.001)), colour = "white", size = 2, alpha = 1) +
  annotate("point", x = head(which(q.seq > q ),1), y = which(adjustments > a12 & adjustments < (a12 + 0.001)), colour = "white", size = 2, alpha = 1) +
  annotate("text", x = head(which(q.seq > q ),1) - 40 , y = which(adjustments > a6 & adjustments < (a6 + 0.001)), label = "6", colour = "white") +
  annotate("text", x = head(which(q.seq > q ),1) -40, y = which(adjustments > a12 & adjustments < (a12 + 0.001)), label = "12", colour = "white") +
theme_bw() +
scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.2")) +
scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
```

#What if we use some other than the default cut-off for model class assingment

```{r}
#Load a data frame with observed classes and probability of deferral from  a classifier
load("../data//rrfFit_roc_validate_probs.rdata")
summary(df)
```

```{r}
#cut prediction.probs with seq(0, 1, .1)
x <- seq(0.1, .9, .1)
pred.classes <- data.frame(matrix(nrow=nrow(df),ncol=length(x)))
colnames(pred.classes) <- paste0("p_",x)
for (i in x ) {
  pred.classes[,paste0("p_",i)] <-  factor( df$Deferred <= i ,labels = c("Deferred","Accepted"))
}

head(pred.classes)
```

```{r}
#count TPR and FPR for each probability
s <- vector()
#df$obs has factor levels in wrong
df$obs <- relevel(df$obs,"Deferred")

for(i in names(pred.classes)) {
  conf <- caret::confusionMatrix(reference=df$obs, data=pred.classes[,i], positive = "Deferred",mode="everything")
  #Sensitivity  = TPR
  s <- c(s,conf$byClass['Sensitivity']) 
}
tprs <- tibble(probability = x,TPR=s,FPR=1-s)
tprs
```

```{r}

get_cost <- function(TPR, FPR,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) {
  q <- d * TPR
  # Find a for 6 month category
  # correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
  # incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
  a6 <- 1 * ((1 - FPR) * (1 - d.)) + 
    3 * mr. * (FPR * (1 - d.)) + 2 * fr. * (FPR * (1 - d.)) + 
    3 * mdr. * d. + 2 * fdr. * d.

  a12 <- 1 * ((1 - FPR) * (1 - d.)) + 
    6 * mr. * (FPR * (1 - d.)) + 4 * fr. * (FPR * (1 - d.)) + 
    6 * mdr. * d. + 4 * fdr. * d.

  # a6
  Fratio <- a6 / (1 + Fn. * (a6 - 1))
  Em <- Pm. * (Fratio - 1 - q * rl.)
  Ed <- Pd. * q
  E6 <- Em - Ed

  # a12
  Fratio <- a12 / (1 + Fn. * (a12 - 1))
  Em <- Pm. * (Fratio - 1 - q * rl.)
  Ed <- Pd. * q
  E12 <- Em - Ed
  tibble(q=q,a6=a6,a12=a12,E6=E6,E12=E12)
}

#Do we get the same as before
# > get_cost(TPR=1662/2116, FPR=1- (1662 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
# # A tibble: 1 x 5
#        q    a6   a12     E6     E12
#    <dbl> <dbl> <dbl>  <dbl>   <dbl>
# 1 0.0257  1.35  1.93 -0.945 -0.0255
#Yes

tprs <- tprs %>% group_by(probability,TPR,FPR) %>%  do(get_cost(.$TPR,.$FPR) )

```

```{r}
p <- ggplot(tprs)
p <- p + geom_line(aes(x=probability,y=E6))
p <- p + geom_line(aes(x=probability,y=E12)) + ylab("cost")
p
```