### 2020-12-01 src/article_figures.Rmd
Use the raw result dataframe as input for performance forest plot.

---
title: "Article figures"
author: "Jarkko Toivonen"
date: "10/30/2020"
output: 
  md_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
save_figs = TRUE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

```{r, include=FALSE}

descript <- tibble(Variable = c("donor", "Hb", "days_to_previous_fb", "age", "previous_Hb_def", 
                                "year", "warm_season", "consecutive_deferrals", "recent_donations",
                                "recent_deferrals", "hour", 
                                "previous_Hb", "Hb_first", "Hb_deferral","nb_donat","gender"), 
                   Pretty = c("Donor ID", "Hemoglobin", "Days to previous full blood donation", "Age", 
                              "Previous donation deferred", #"Previous Hb deferral",   # This is a more descriptive name
                              "Year", "Warm season", "Consecutive deferrals", "Recent donations", 
                              "Recent deferrals", "Hour", 
                              "Previous Hb", "First Hb", "Hb deferral","Life time donations","Sex"),
                   Type = c("Factor", "numeric", "numeric (int)", "numeric", "boolean",
                            "numeric (int)", "boolean", "numeric (int)", "numeric (int)", "numeric (int)", "numeric",
                            "numeric", "numeric", "boolean","numeric (int)","Factor"),
                   Explanation = c("Donor identifier",
                                   "Amount of Hemoglobin",
                                   "Time (in days) between Hb measurement and previous full blood donation event",
                                   "Age of donor",
                                   "Indicates whether the donor was deferred from blood donation due to low hemoglobin at previous donation event",
                                   "Year of donation",
                                   "True if donation was given in April-September",
                                   "Amount of times the donor has been deferred due to low hemoglobin since last succesful whole blood donation",
                                   "Amount of donations in the last two years",
                                   "Amount of deferrals due to low hemoglobin in the last two years",
                                   "Time of day when donation was given as hours (e.g. 13:45 = 13.75)",
                                   "Hb value at previous measurement (ICP-model)",
                                   "Hb value at first donation of this donor (Non ICP-model)",
                                   "Deferred based on low hemoglogin",
                                  "Life time total count of donations",
                                  "Sex"
                   )
                   )

```

## Effect sizes and importances of variables

Effect sizes of variables from eProgesa and Biobank using dynamic linear mixed model. In addition, variables
importance from random forest algorithm.

```{r, include=FALSE}

get_dlmm_df <- function(male_posterior, female_posterior, variables, combined_variables) {
  for (gender in c("male", "female")) {
    #posterior <- ifelse(gender == "male", male_posterior, female_posterior)
    if (gender == "male") {
      posterior <- male_posterior
    } else {
      posterior <- female_posterior
    }

    for (i in seq_along(posterior)) {
      #cat(sprintf("Gender %s, Column %i\n", gender, i))
      v <- posterior[[i]]
      ci_hdi <- bayestestR::ci(v, method = "HDI", ci=0.95)
      #str(ci_hdi)
      L <- list(names = variables[[i]], gender=gender, mean = mean(v), low = ci_hdi$CI_low, high = ci_hdi$CI_high)
      if (i == 1 && gender=="male") {
        result <- data.frame(L, stringsAsFactors = FALSE)
      } else {
        result <- rbind(result, L)
      }
    }
  }
  result <- as_tibble(result)
  result <- result %>%
    mutate(gender = factor(gender, levels=c("female", "both", "male"))) %>%
    rename(Pretty=names) %>%
    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
  result <- left_join(combined_variables, result, by="Pretty")
  return(result)

}

get_rf_df <- function(rrfFit_roc, descript, combined_variables) {
  rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
  #rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
  rrfFit.varimp <- rownames_to_column(rrfFit_rocImp$importance)
  colnames(rrfFit.varimp) <- c("Variable","Importance")
  rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable, Pretty, Importance) %>% arrange(Importance)

  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"

  rrfFit.varimp <- rrfFit.varimp %>% 
    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)),
           gender=factor("both", levels=c("female", "both", "male")))
  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")
  return(rrfFit.varimp)  
}

# Creates a forest plot of means and confidence intervals from posterior distributions.
# Puts both male and female results in the same plot.
create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  cis <- result

  result <- result %>%
    mutate(Pretty=fct_rev(Pretty))

  plot <- result %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    geom_rect(data=combined_variables %>% filter(stripe==1), 
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                  ymin = as.numeric(Pretty) - 0.5),
                  #xmin = -Inf,
                  #xmax = Inf,
                  fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    labs(title="", #title="Effects sizes of variables on Hb prediction",
         x="Regression coefficient", y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
    #theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
  return(list(plot=plot, cis=cis))
}

create_variable_importance_plot2 <- function(rrfFit_roc, descript, combined_variables, base_size = 11) {
  rrfFit.varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)

  var_imp <- rrfFit.varimp %>%
      mutate(Pretty=fct_rev(Pretty))

  varimp.plot <- var_imp %>% 
    ggplot() + 
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              #xmin = -Inf,
              #xmax = Inf,
              #fill = stripe), 
              xmin=-Inf, xmax=Inf, alpha = 0.1, colour=NA, fill="gray", show.legend = FALSE, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
                               #colour = hue_pal()(3)[2], size=0.25) +
    labs(y=NULL, colour="Sex") + 
    guides(y="none") +
    scale_colour_discrete(drop=FALSE, breaks=c("both")) +
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    #theme_gray(base_size = base_size)
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")

  return(varimp.plot)
}

create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
    mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
           Pretty=factor(Pretty, levels=Pretty))

  forest <- create_double_forest_plot(male_posterior, female_posterior, variables, combined_variables, base_size = base_size)$plot
  importance <- create_variable_importance_plot2(rrfFit_roc, descript, combined_variables, base_size = base_size)

  use_cowplot <- TRUE

  if (use_cowplot) {
    forest_importance <- cowplot::plot_grid(forest, importance, labels = c('A', 'B'), label_size = 12, nrow=1, scale=1.0, axis="tb", align="h",
                                            rel_widths = c(2, 1))
    if (!is.null(filename)) {
      cowplot::save_plot(filename, forest_importance, title="Effect sizes and importances",
                         ncol = 2, base_asp = 1.0, base_width = width / 25.4 / 2, base_height = NULL)
    }
  } else {
    forest_importance <- gridExtra::grid.arrange(forest, importance, nrow = 1, respect=TRUE)   # Combine the plots
    if (!is.null(filename)) {
      ggsave(filename=filename, forest_importance, width = width, units="mm", dpi=600, scale=1.0)
    }
  }
  forest_importance
}
```

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
                    `Previous Hb deferral` = "Previous donation deferred",
                    `Polygenic risk score` = "Polygenic score",
                    `RNF43 mutant` = "RNF43")
```

```{r, eval=FALSE}
if (save_figs) {
    filename <- paste(fig_path, "old-effect-size-importance.pdf", sep="/")
} else {
    filename <- NULL
}
create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
```

## Alternative effect sizes and importances of variables

```{r, include=FALSE}
alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
    mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
           Pretty=factor(Pretty, levels=Pretty))

  # Get effect sizes of the linear model  
  dlmm <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
  dlmm <- dlmm %>%
    mutate(Pretty=fct_rev(Pretty))

  # Get variable importance from the RF object
  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=ifelse(is.na(Importance), NA, 0),
           high=Importance)

  final <- bind_rows(`Regression coefficient`=dlmm, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
    select(-Importance, -Variable, stripe)

  print(combined_variables)

  plot <- final %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), 
                               size=0.25,  
                               inherit.aes = FALSE) +
    labs(title=NULL, x=NULL, y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) + 
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    facet_wrap("method", scales="free_x") +
    theme_classic(base_size = base_size) + 
    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))#,
          #plot.background = element_rect(fill = "lightblue"))

  return(list(plot=plot, final=final))
}
```

```{r}
res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
g <- res$plot
final <- res$final
if (save_figs)
  ggsave(filename="effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180)
  ggsave(filename="presentation-effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180, height=90)
g
```

## Performance forest plot

```{r, include=FALSE}

all_ids <- c(ids, dummy2_ids)

get_threshold_df <- function() {

  t6 <- tibble(
      Id = factor(all_ids),
      type = factor("Threshold"),
      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6,   0.2, 0.3))
  t12 <- tibble(
      Id = factor(all_ids),
      type = factor("Threshold"),
      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8,   0.4, 0.5)
  )
  t <- bind_rows(t6, t12)
  t <- t %>%
    mutate(low=value, 
           high=value,
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           #model=factor(str_split_fixed(Id, "-", 3)[,3])
           model=rep(c("t6", "t12"), each=length(all_ids))
           )
  return(t)
}

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
  # s <- map(s, function(x) when(x, 
  #                              x == "progesa" ~ "eProgesa",
  #                              x == "finngen" ~ "Biobank",
  #                              x == "findonor" ~ "FinDonor",
  #                              ~ "koe"))
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
  if (FALSE) {
    df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
    df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
    df <- bind_rows(df1, df2)
    #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
      mutate(type = fct_recode(type, !!!new_names))

    thresholds <- get_threshold_df()
    df <- bind_rows(df, thresholds)
    df <- df %>% 
      mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
  } else {
#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")

    df <- df %>% 
      pivot_wider(names_from=c("type"), values_from=value) %>%
      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
             high = ifelse(is.na(high), value, high))

    df <- df %>%
      mutate(Id=factor(Id, levels=rev(all_ids)),
             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
             model=case_when(variable == "threshold6" ~ "t6",
                             variable == "threshold12" ~ "t12",
                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")),
             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
             variable = fct_recode(variable, !!!new_names))

  }

  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(variable %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, variable=mytype) # Dummy table to use the same x-axis limits for E6 and E12

  dummy <- dummy %>%
    mutate(variable = fct_recode(variable, !!!new_names))
  print(dummy)
  print(df)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Dummy"),
                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("variable", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  return(list(g=g, df=df))
}

```

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs)
  ggsave(filename="performance-forest-plot.pdf", #  title="Performance forest plot", 
         path=fig_path, plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
g
```

## Calibration plots

```{r calibration plots, include=FALSE}
calibration_plot <- function(df, use_my_method=FALSE) {
  if (use_my_method) {
    df$cat <- cut(df$Deferred, seq(0.0, 1.0, length.out = 11))
    df2 <- df %>% group_by(cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"))
    df2 %>% ggplot(aes(cat, true_bin_prob)) + geom_point() + ylim(0, 1)
  } else {
    cal <- calibration(obs ~ Deferred, data=df, class="Deferred", cuts=11)
    ggplot(cal) + ylim(0, 100)
  }
}

calibration_plots <- function(ids) {
  data_frames <- map(ids, get_data_frame)
  names(data_frames) <- ids
  df <- bind_rows(data_frames, .id="Id")
  df <- df %>% 
    group_by(Id) %>%
    mutate(cat=cut(Deferred, seq(0.0, 1.0, length.out = 11))) %>%
    group_by(Id, cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"), count=n()) %>%
    ungroup() %>%
    mutate(Id = factor(Id, levels=ids))

  g <- df %>% ggplot(aes(cat, true_bin_prob, colour=count)) + 
    geom_point() + 
    geom_text(aes(label=count), colour="black", nudge_y=0.05) +
    ylim(0, 1.05) +
    labs(x="Predicted probability of deferral", y="True probability of deferral in each bin") +
    scale_colour_gradient(name = "Count", trans = "log10") +
    geom_abline(aes(intercept=-0.1, slope=0.1)) +
    facet_wrap("Id") +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

  #return(df)
  return(g)
}
```

```{r cars}
g <- calibration_plots(ids)
if (save_figs)
  ggsave(filename="calibration_plots.pdf", title="Calibration plots", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

## Classification scatter plot for male Finngen DLMM

```{r, include=FALSE}

generate_my_breaks <- function(step) {
  # Break between limits at every position that is multiple of 'step' 
  my_breaks <- function(limits) {
    #step <- 0.2
    m <- limits %/% step
    m <- ifelse(m < 0, m+1, m)
    m <- m*step
    return(seq(m[1], m[2], step))
  }
  return(my_breaks)
}

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
    df <- df %>% 
      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
  } else { # predict deferral based on probability of deferral
    df <- df %>% 
      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
  }
  df <- df %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))
  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_colour_discrete(drop=FALSE) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = hb_threshold, linetype = "dashed") +
    theme(legend.position = "bottom")
    #ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")

```

```{r}
g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g2, dpi=600, units="mm", width=180)
}
g1
g2
```

Testing the effect of base_size parameter.

```{r}
base_size <- 20
presentation1 <- g1 + coord_fixed() + theme_gray(base_size=base_size)
presentation2 <- g2 + coord_fixed() + theme_gray(base_size=base_size)
if (save_figs) {
  ggsave(filename="presentation-classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=presentation1, dpi=600, units="mm", width=180, height=120)
  ggsave(filename="presentation-classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=presentation2, dpi=600, units="mm", width=180, height=120)
}

```

Pretty print results as a table.

```{r}
df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
df <- df %>% 
  pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
names(df) <- str_remove(names(df), " value$")
df %>%
  select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
```

```{r, eval=FALSE, include=FALSE}
# jarkko_test <- function(male_posterior, female_posterior, variables,
#                                            filename=NULL,
#                                            width = 180,  # width of the combined figure in millimetres
#                                           base_size = 11
# ) {
#   rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
# 
#   combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
#     mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
#            Pretty=factor(Pretty, levels=Pretty))
# 
#   result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
#   varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
# 
#   result <- result %>%
#     mutate(Pretty=fct_rev(Pretty))
# 
#   var_imp <- varimp %>%
#     mutate(Pretty=fct_rev(Pretty),
#            mean=Importance,
#            low=ifelse(is.na(Importance), NA, 0),
#            high=Importance)
# 
#   final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
#     mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
#     select(-Importance, -Variable, stripe)
# 
# #   final <- tibble(
# #             method = c("Regression coefficient",
# #                        "Regression coefficient","Importance"),
# #             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
# #                        "Days to previous full blood donation")),
# #             #gender = c("male", "female", "both"),
# #               mean = c(0.0152091681221149,
# #                        0.0445741103338152,0.0255604042298423),
# #                low = c(0.00691102953212081, 0.0353192308283623, 0),
# #               high = c(0.0240855557569598,
# #                        0.0537929726515129,0.0255604042298423)
# # )
#   
#   #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))
#   
#   print(final)         
#   print(combined_variables)
#   vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
#   plot <- final %>% #filter(Pretty %in% vars) %>%
#     ggplot(aes(color=gender)) +
#     geom_rect(data=combined_variables %>% filter(stripe==1),
#               mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
#                           #ymin = as.numeric(Pretty) - 0.5
#                           ymax = as.numeric(Pretty) + 0.5,
#                           ymin = as.numeric(Pretty) - 0.5),
#               fill = "blue", 
#               #position="identity",
#               xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
#     ggstance::geom_pointrangeh(#data=final, 
#                                mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
#                                #position=position_dodge2(0.8)
#                                position=position_dodge2(width=0.8, padding=0.1)
#                                ) +
#     scale_y_discrete() +
#     facet_wrap("method", scales = "free_x") +
#     #coord_flip() +
#     theme_classic(base_size = base_size) 
# 
#   return(list(plot=plot, final=final))
# }
# 
# res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
# final <- res$final
# g <- res$plot
# g
```